Temporal Dynamics of Normalization Reweighting

### 2023-07-31 manuscript.Rmd
Implemented reviewer suggestions

---
title: 'Temporal Dynamics of Normalization Reweighting'
author: | 
  | Daniel H. Baker, Daniela Marinova, Richard Aveyard, Lydia J. Hargreaves, Alice Renton,  
  | Ruby Castellani, Phoebe Hall, Miriam Harmens, Georgia Holroyd, Beth Nicholson, 
  | Emily L. Williams, Hannah M. Hobson & Alex R. Wade
date: "Department of Psychology, University of York, UK"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: yes
    extra_dependencies: ["lineno"]
  bookdown::word_document2:
    fig_caption: yes
    toc: no
  word_document: default
  html_document: default
bibliography: references.bib
csl: elife.csl
---

@@

plotsmoothed <- 1 # either plot the raw data (0), or a smoothed spline (1)
useSPQ <- 0  # optionally do the median split by SPQ for the main EEG experiment

@@

```{r include=FALSE, results='hide'}

@@

samplerate <- 1000
targetF <- 7
maskF <- 5
tindex <- (10*targetF) + 1
mindex <- (10*maskF) + 1
duration <- 14

@@

```{r include=FALSE, results='hide'}

@@

    spec <- fft(temp[2501:12500])/10000
    EEGspectra[s,cond,] <- spec[1:(20*10)]

for (t in 1:(14000-binwidth)){
  spec <- fft(temp[t:(t+binwidth-1)])/binwidth
  for (f in 1:length(findex)){
  EEGtimecourse[s,f,cond,t] <- spec[findex[f]]
  }
}
  }

@@

```{r include=FALSE, results='hide'}

@@

spl5a <- smooth.spline(times,EEGmeantimes[1,1,],df=20)
lines(spl5a$x,spl5a$y,col=colpal[1],lwd=2)
spl5a <- smooth.spline(times,EEGmeantimes[1,2,],df=20)
lines(spl5a$x,spl5a$y,col='grey',lwd=2)
spl5a <- smooth.spline(times,EEGmeantimes[1,3,],df=20)
lines(spl5a$x,spl5a$y,col=colpal[2],lwd=2)
}
text(10,0.95,'7Hz',cex=2)

@@

spl5aU <- smooth.spline(times,EEGmeantimes[2,2,]+EEGSEtimes[2,2,],df=20)
spl5aL <- smooth.spline(times,EEGmeantimes[2,2,]-EEGSEtimes[2,2,],df=20)
polygon(times[c(1:13000,13000:1)],c(spl5aU$y,spl5aL$y[13000:1]),col=colpal[1],border=NA)

@@

  pdf('Figures/VilidaiteData.pdf', bg="transparent", height = 6, width = 16)
  par(mar=c(0.1,0.1,0.1,0.1))
  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,1), ylim=c(0,1))

  aspratio <- 6/16  # this is the aspect ratio of the output pdf
  imwidth <- 0.25
  xstart <- 0.02
  ystart <- 0.58
  rasterImage(e1,xstart,ystart,xstart+imwidth*aspratio,ystart+imwidth)
  xstart <- 0.22
  ystart <- 0.58
  rasterImage(e2,xstart,ystart,xstart+imwidth*aspratio,ystart+imwidth)

  grid.picture(e3,x=0.225,y=0.7,width=0.4,height=1)
  grid.picture(e5,x=0.225,y=0.25,width=0.4,height=1)

  grid.picture(e6,x=0.5,y=0.85,width=0.1,height=1)
  grid.picture(e7,x=0.5,y=0.625,width=0.1,height=1)
  grid.picture(e8,x=0.5,y=0.375,width=0.1,height=1)
  grid.picture(e9,x=0.5,y=0.15,width=0.1,height=1)
  grid.picture(e4,x=0.77,y=0.5,width=0.45,height=1)

  text(-0.02,0.98,'(a)',pos=4,cex=2)
  text(-0.02,0.45,'(b)',pos=4,cex=2)
  text(0.41,0.98,'(c)',pos=4,cex=2)
  text(0.57,0.98,'(d)',pos=4,cex=2)

@@

```{r include=FALSE, results='hide'}

@@

@@

```{r include=FALSE, results='hide'}

@@

```{r include=FALSE, results='hide'}

@@

fspec1 <- NULL
fspec2 <- NULL
for (t in 1:(9000-binwidth)){
  spec <- fft(temp[t:(t+binwidth-1)])/binwidth
  fspec1[t] <- spec[tindex]
  fspec2[t] <- spec[mindex]
}
    EEGtimecourse5Hz[s,cond,] <- fspec1
    EEGtimecourse7Hz[s,cond,] <- fspec2

@@

```{r include=FALSE, results='hide'}

@@

isincluded <- matrix(rep(1,4*dim(allspec)[1]),nrow=4,ncol=dim(allspec)[1])
for (cond in 1:4){
  # note that in allspec the DC component has been removed
  # so we don't need to add 1 to the index
      tempxy <- data.frame(Re(allspec[,cond,duration*targetF]),Im(allspec[,cond,duration*targetF]))
      D5 <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))
      tempxy <- data.frame(Re(allspec[,cond,duration*maskF]),Im(allspec[,cond,duration*maskF]))      
      D7 <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))      
      isincluded[cond,which(D5>3)] <- 0
      isincluded[cond,which(D7>3)] <- 0
   for (s in 1:100){
     if (max(abs(EEGtimecourse5Hz[s,cond,]))>5){isincluded[cond,s] <- 0}
   }   
}

@@

EEGmaskindex5 <- matrix(0,nrow=2,ncol=8000)
EEGmaskindexSE5 <- matrix(0,nrow=2,ncol=8000)
includedboth <- isincluded[1,]*isincluded[3,]
monratios <- abs(EEGtimecourse5Hz[which(includedboth==1),1,])/abs(EEGtimecourse5Hz[which(includedboth==1),3,])
EEGmaskindex5[1,] <- colMeans(20*log10(monratios))
EEGmaskindexSE5[1,] <- apply(20*log10(monratios),2,sd)/sqrt(sum(includedboth))
includedboth <- isincluded[1,]*isincluded[4,]
dichratios <- abs(EEGtimecourse5Hz[which(includedboth==1),1,])/abs(EEGtimecourse5Hz[which(includedboth==1),4,])
EEGmaskindex5[2,] <- colMeans(20*log10(dichratios))
EEGmaskindexSE5[2,] <- apply(20*log10(dichratios),2,sd)/sqrt(sum(includedboth))

@@

@@

@@

EEGmaskindex7 <- matrix(0,nrow=2,ncol=8000)
EEGmaskindexSE7 <- matrix(0,nrow=2,ncol=8000)
includedboth <- isincluded[2,]*isincluded[3,]
monratios <- abs(EEGtimecourse7Hz[which(includedboth==1),2,])/abs(EEGtimecourse7Hz[which(includedboth==1),3,])
EEGmaskindex7[1,] <- colMeans(20*log10(monratios))
EEGmaskindexSE7[1,] <- apply(20*log10(monratios),2,sd)/sqrt(sum(includedboth))
includedboth <- isincluded[2,]*isincluded[4,]
dichratios <- abs(EEGtimecourse7Hz[which(includedboth==1),2,])/abs(EEGtimecourse7Hz[which(includedboth==1),4,])
EEGmaskindex7[2,] <- colMeans(20*log10(dichratios))
EEGmaskindexSE7[2,] <- apply(20*log10(dichratios),2,sd)/sqrt(sum(includedboth))

@@

@@

if (plotsmoothed==0){
polygon(times[c(1:8000,8000:1)],c(EEGmeantime5Hz[1,]+EEGSEtime5Hz[1,],EEGmeantime5Hz[1,8000:1]-EEGSEtime5Hz[1,8000:1]),col=colpal[1],border=NA)
polygon(times[c(1:8000,8000:1)],c(EEGmeantime5Hz[3,]+EEGSEtime5Hz[3,],EEGmeantime5Hz[3,8000:1]-EEGSEtime5Hz[3,8000:1]),col=colpal[2],border=NA)
polygon(times[c(1:8000,8000:1)],c(EEGmeantime5Hz[4,]+EEGSEtime5Hz[4,],EEGmeantime5Hz[4,8000:1]-EEGSEtime5Hz[4,8000:1]),col=colpal[3],border=NA)

lines(times,EEGmeantime5Hz[1,],lwd=2,col=colpal[1])
# lines(times,EEGmeantime5Hz[2,],lwd=2,col='green')
lines(times,EEGmeantime5Hz[3,],lwd=2,col=colpal[2])
lines(times,EEGmeantime5Hz[4,],lwd=2,col=colpal[3])
}

if (plotsmoothed==1){
spl5aU <- smooth.spline(times,EEGmeantime5Hz[1,]+EEGSEtime5Hz[1,],df=20)
spl5aL <- smooth.spline(times,EEGmeantime5Hz[1,]-EEGSEtime5Hz[1,],df=20)
polygon(times[c(1:8000,8000:1)],c(spl5aU$y,spl5aL$y[8000:1]),col=colpal[1],border=NA)

@@

spl5a <- smooth.spline(times,EEGmeantime5Hz[1,],df=20)
lines(spl5a$x,spl5a$y,col=colpal[1],lwd=2)
# spl5b <- smooth.spline(times,EEGmeantime5Hz[2,],df=20)
# lines(spl5b$x,spl5b$y,col='green',lwd=2)
spl5c <- smooth.spline(times,EEGmeantime5Hz[3,],df=20)
lines(spl5c$x,spl5c$y,col=colpal[2],lwd=2)
spl5d <- smooth.spline(times,EEGmeantime5Hz[4,],df=20)
lines(spl5d$x,spl5d$y,col=colpal[3],lwd=2)

}
text(6,0.55,'5Hz',cex=2)
  legend(1.25,0.2,c('Baseline','Monocular','Dichoptic'),lty=1,lwd=2,col=colpal[1:3],box.lwd=2)
dev.off()

@@

  pdf('Figures/EEGdata.pdf', bg="transparent", height = 6, width = 15)
  par(mar=c(0.1,0.1,0.1,0.1))
  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,1), ylim=c(0,1))

@@

```{r include=FALSE, results='hide'}

@@

@@

```{r include=FALSE, results='hide'}

@@

        fspec1 <- NULL
        fspec2 <- NULL
        for (t in 1:(9009-binwidth)){
          spec <- fft(temp[t:(t+binwidth-1)])/binwidth
          fspec1[t] <- spec[tindex]
          fspec2[t] <- spec[mindex]
        }

        MEGtimecourse5Hz[s,cond,] <- fspec1
        MEGtimecourse7Hz[s,cond,] <- fspec2

@@

```{r include=FALSE, results='hide'}

@@

  isincluded <- array(1, dim=c(4,dim(allspec)[1]))
    for (cond in 1:4){
      # note that in allspec the DC component has been removed
      # so we don't need to add 1 to the index
      tempxy <- data.frame(Re(allspec[,cond,duration*targetF]),Im(allspec[,cond,duration*targetF]))
      D5 <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))
      tempxy <- data.frame(Re(allspec[,cond,duration*maskF]),Im(allspec[,cond,duration*maskF]))      
      D7 <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))      
      isincluded[cond,which(D5>3)] <- 0
      isincluded[cond,which(D7>3)] <- 0
      for (s in 1:20){
        if (max(abs(MEGtimecourse5Hz[s,cond,]))>5){isincluded[cond,s] <- 0}
      }   
    }

@@

    includedboth <- isincluded[1,]*isincluded[3,]
    monratios <- abs(MEGtimecourse5Hz[which(includedboth==1),1,])/abs(MEGtimecourse5Hz[which(includedboth==1),3,])
    MEGmaskindex5[1,] <- colMeans(20*log10(monratios))
    MEGmaskindexSE5[1,] <- apply(20*log10(monratios),2,sd)/sqrt(sum(includedboth))
    includedboth <- isincluded[1,]*isincluded[4,]
    dichratios <- abs(MEGtimecourse5Hz[which(includedboth==1),1,])/abs(MEGtimecourse5Hz[which(includedboth==1),4,])
    MEGmaskindex5[2,] <- colMeans(20*log10(dichratios))
    MEGmaskindexSE5[2,] <- apply(20*log10(dichratios),2,sd)/sqrt(sum(includedboth))

@@

@@

  }
  text(6,2.8,'5Hz',cex=2)
  legend(1.25,1,c('Baseline','Monocular','Dichoptic'),lty=1,lwd=2,col=colpal[1:3],box.lwd=2)
  dev.off()

@@

  pdf('Figures/MEGdata.pdf', bg="transparent", height = 6, width = 15)
  par(mar=c(0.1,0.1,0.1,0.1))
  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,1), ylim=c(0,1))

@@

@@

```{r include=FALSE, results='hide'}

>>>>

---
title: 'Temporal Dynamics of Normalization Reweighting'
author: | 
  | Daniel H. Baker, Daniela Marinova, Richard Aveyard, Lydia J. Hargreaves, Alice Renton,
  | Ruby Castellani, Phoebe Hall, Miriam Harmens, Georgia Holroyd, Beth Nicholson, 
  | Emily L. Williams, Hannah M. Hobson & Alex R. Wade
date: "Department of Psychology, University of York, UK"
output:
  bookdown::pdf_document2:
    fig_caption: yes
    toc: no
    keep_tex: yes
    extra_dependencies: ["lineno"]
bibliography: references.bib
link-citations: true
csl: journal-of-vision.csl
---

@@

plotsmoothed <- 1 # either plot the raw data (0), or a smoothed spline (1)
fixedphase <- 0 # option to use a fixed phase sinusoidal dot product instead of Fourier analysis for the timecourses
useSPQ <- 0  # optionally do the median split by SPQ for the main EEG experiment

@@

```{r analysepilotindividual, include=FALSE, results='hide'}

@@

samplerate <- 1000
targetF <- 7
maskF <- 5
tindex <- (10*targetF) + 1
mindex <- (10*maskF) + 1
tindex2 <- (10*targetF*2) + 1
mindex2 <- (10*maskF*2) + 1
duration <- 14

@@

```{r analysepilotgroup, include=FALSE, results='hide'}

@@

    spec <- fft(temp[2501:12500])/10000
    EEGspectra[s,cond,] <- spec[1:(20*10)]
if (fixedphase==0){  # do normal Fourier analysis
for (t in 1:(14000-binwidth)){
  spec <- fft(temp[t:(t+binwidth-1)])/binwidth
  for (f in 1:length(findex)){
  EEGtimecourse[s,f,cond,t] <- spec[findex[f]]
  }
}}

if (fixedphase==1){  # used fixed phase dot product with sinewave
  # make a sine wave with the same phase as the EEG response
  spec <- fft(temp)/length(temp)
  synspec <- spec*0
  synspec[(targetF*14) + 1] <- spec[(targetF*14) + 1]
  sw1 <- Re(fft(synspec,inverse=TRUE))
  sw1 <- sw1/max(sw1)
  synspec <- spec*0
  synspec[(maskF*14) + 1] <- spec[(maskF*14) + 1]
  sw2 <- Re(fft(synspec,inverse=TRUE))  
  sw2 <- sw2/max(sw2)
  synspec <- spec*0
  synspec[(2*targetF*14) + 1] <- spec[(2*targetF*14) + 1]
  sw3 <- Re(fft(synspec,inverse=TRUE))
  sw3 <- sw3/max(sw3)
  synspec <- spec*0
  synspec[(2*maskF*14) + 1] <- spec[(2*maskF*14) + 1]
  sw4 <- Re(fft(synspec,inverse=TRUE)) 
  sw4 <- sw4/max(sw4)

for (t in 1:(14000-binwidth)){
  EEGtimecourse[s,1,cond,t] <- mean(temp[t:(t+binwidth-1)] * sw1[t:(t+binwidth-1)]) + 0.1
  EEGtimecourse[s,2,cond,t] <- mean(temp[t:(t+binwidth-1)] * sw2[t:(t+binwidth-1)]) + 0.1
  EEGtimecourse[s,3,cond,t] <- mean(temp[t:(t+binwidth-1)] * sw3[t:(t+binwidth-1)]) + 0.1
  EEGtimecourse[s,4,cond,t] <- mean(temp[t:(t+binwidth-1)] * sw4[t:(t+binwidth-1)]) + 0.1
}}

@@

```{r plotpilotdata, include=FALSE, results='hide'}

@@

spl5a <- smooth.spline(times,EEGmeantimes[1,1,],df=20)
lines(spl5a$x,spl5a$y,col=colpal[1],lwd=2)
spl5a <- smooth.spline(times,EEGmeantimes[1,2,],df=20)
lines(spl5a$x,spl5a$y,col='grey',lwd=2)
spl5a <- smooth.spline(times,EEGmeantimes[1,3,],df=20)
lines(spl5a$x,spl5a$y,col=colpal[2],lwd=2)
}

legend(2,1,c('Baseline','Target','Target+mask'),col=c('grey',colpal[1],colpal[2]),lty=1,lwd=2,box.lwd=2)

text(10,0.95,'7Hz',cex=2)

@@

spl5aU <- smooth.spline(times,EEGmeantimes[2,2,]+EEGSEtimes[2,2,],df=20)
spl5aL <- smooth.spline(times,EEGmeantimes[2,2,]-EEGSEtimes[2,2,],df=20)
 # use of pmin function here crops the error region at the top of the graph 
polygon(times[c(1:13000,13000:1)],c(pmin(spl5aU$y,1),spl5aL$y[13000:1]),col=colpal[1],border=NA)

@@

 if (fixedphase==0){pdf('Figures/VilidaiteData.pdf', bg="transparent", height = 15, width = 6)}
 if (fixedphase==1){pdf('Figures/VilidaiteDataFP.pdf', bg="transparent", height = 15, width = 6)}  
  par(mar=c(0.1,0.1,0.1,0.1))
  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,1), ylim=c(0,1))

  aspratio <- 15/6  # this is the aspect ratio of the output pdf
  imwidth <- 0.05
  xstart <- 0.05
  ystart <- 0.92
  rasterImage(e1,xstart,ystart,xstart+imwidth*aspratio,ystart+imwidth)
  xstart <- 0.55
  ystart <- 0.92
  rasterImage(e2,xstart,ystart,xstart+imwidth*aspratio,ystart+imwidth)

  grid.picture(e3,x=0.5,y=0.91,width=1)
  grid.picture(e5,x=0.5,y=0.74,width=1)

  grid.picture(e6,x=0.25,y=0.57,width=0.5,height=1)
  grid.picture(e7,x=0.75,y=0.57,width=0.5,height=1)
  grid.picture(e8,x=0.25,y=0.42,width=0.5,height=1)
  grid.picture(e9,x=0.75,y=0.42,width=0.5,height=1)

  grid.picture(e4,x=0.5,y=2/12,width=1)

  text(-0.055,1.02,'(a)',pos=4,cex=2)
  text(-0.055,0.85,'(b)',pos=4,cex=2)
  text(-0.055,0.67,'(c)',pos=4,cex=2)
  text(-0.055,0.3,'(d)',pos=4,cex=2)

@@

```{r pilotstats, include=FALSE, results='hide'}

@@

# calculate effect size

baseline <- rowMeans((20*log10(abs(EEGtimecourse[,1,1,1001:2000])/abs(EEGtimecourse[,1,3,1001:2000])) + 20*log10(abs(EEGtimecourse[,2,2,1001:2000])/abs(EEGtimecourse[,2,3,1001:2000])) + 20*log10(abs(EEGtimecourse[,3,1,1001:2000])/abs(EEGtimecourse[,3,3,1001:2000])) + 20*log10(abs(EEGtimecourse[,4,2,1001:2000])/abs(EEGtimecourse[,4,3,1001:2000])))/4)

laterepoch <- rowMeans((20*log10(abs(EEGtimecourse[,1,1,3001:4000])/abs(EEGtimecourse[,1,3,3001:4000])) + 20*log10(abs(EEGtimecourse[,2,2,3001:4000])/abs(EEGtimecourse[,2,3,3001:4000])) + 20*log10(abs(EEGtimecourse[,3,1,3001:4000])/abs(EEGtimecourse[,3,3,3001:4000])) + 20*log10(abs(EEGtimecourse[,4,2,3001:4000])/abs(EEGtimecourse[,4,3,3001:4000])))/4)

diff <- laterepoch-baseline

piloteffectsize <- mean(diff)/sd(diff)

@@

```{r analyseEEGindividual, include=FALSE, results='hide'}

@@

```{r analyseEEGgroup, include=FALSE, results='hide'}

@@

if (fixedphase==0){
fspec1 <- NULL
fspec2 <- NULL
for (t in 1:(9000-binwidth)){
  spec <- fft(temp[t:(t+binwidth-1)])/binwidth
  fspec1[t] <- spec[tindex]
  fspec2[t] <- spec[mindex]
}
    EEGtimecourse5Hz[s,cond,] <- fspec1
    EEGtimecourse7Hz[s,cond,] <- fspec2
}

if (fixedphase==1){  # used fixed phase dot product with sinewave
  # make a sine wave with the same phase as the EEG response
  spec <- fft(temp)/length(temp)
  synspec <- spec*0
  synspec[(targetF*9) + 1] <- spec[(targetF*9) + 1]
  sw1 <- Re(fft(synspec,inverse=TRUE))
  sw1 <- sw1/max(sw1)
  synspec <- spec*0
  synspec[(maskF*9) + 1] <- spec[(maskF*9) + 1]
  sw2 <- Re(fft(synspec,inverse=TRUE)) 
  sw2 <- sw2/max(sw2)

for (t in 1:(9000-binwidth)){
  EEGtimecourse5Hz[s,cond,t] <- mean(temp[t:(t+binwidth-1)] * sw1[t:(t+binwidth-1)])
  EEGtimecourse7Hz[s,cond,t] <- mean(temp[t:(t+binwidth-1)] * sw2[t:(t+binwidth-1)])
}}

@@

```{r plotEEGdata, include=FALSE, results='hide'}

@@

isincluded <- matrix(rep(1,4*dim(allspec)[1]),nrow=4,ncol=dim(allspec)[1])
if (fixedphase==0){
for (cond in 1:4){
  # note that in allspec the DC component has been removed
  # so we don't need to add 1 to the index
      tempxy <- data.frame(Re(allspec[,cond,duration*targetF]),Im(allspec[,cond,duration*targetF]))
      D5 <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))
      tempxy <- data.frame(Re(allspec[,cond,duration*maskF]),Im(allspec[,cond,duration*maskF]))      
      D7 <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))      
      isincluded[cond,which(D5>3)] <- 0
      isincluded[cond,which(D7>3)] <- 0
   for (s in 1:100){
     if (max(abs(EEGtimecourse5Hz[s,cond,]))>5){isincluded[cond,s] <- 0}
   }   
}
}

@@

EEGmaskindex5 <- matrix(0,nrow=2,ncol=8000)
EEGmaskindexSE5 <- matrix(0,nrow=2,ncol=8000)
includedboth <- isincluded[1,]*isincluded[3,]
monratios <- abs(EEGtimecourse5Hz[which(includedboth==1),1,])/abs(EEGtimecourse5Hz[which(includedboth==1),3,])
monratios[which(monratios==0)] <- 0.001
EEGmaskindex5[1,] <- colMeans(20*log10(monratios),na.rm=TRUE)
EEGmaskindexSE5[1,] <- apply(20*log10(monratios),2,sd,na.rm=TRUE)/sqrt(sum(includedboth))
includedboth <- isincluded[1,]*isincluded[4,]
dichratios <- abs(EEGtimecourse5Hz[which(includedboth==1),1,])/abs(EEGtimecourse5Hz[which(includedboth==1),4,])
dichratios[which(dichratios==0)] <- 0.001
EEGmaskindex5[2,] <- colMeans(20*log10(dichratios),na.rm=TRUE)
EEGmaskindexSE5[2,] <- apply(20*log10(dichratios),2,sd,na.rm=TRUE)/sqrt(sum(includedboth))

@@

earlydataM[is.na(earlydataM)] <- 0

@@

earlydataD[is.na(earlydataD)] <- 0

@@

EEGmaskindex7 <- matrix(0,nrow=2,ncol=8000)
EEGmaskindexSE7 <- matrix(0,nrow=2,ncol=8000)
includedboth <- isincluded[2,]*isincluded[3,]
monratios <- abs(EEGtimecourse7Hz[which(includedboth==1),2,])/abs(EEGtimecourse7Hz[which(includedboth==1),3,])
monratios[which(monratios==0)] <- 0.001
EEGmaskindex7[1,] <- colMeans(20*log10(monratios),na.rm=TRUE)
EEGmaskindexSE7[1,] <- apply(20*log10(monratios),2,sd,na.rm=TRUE)/sqrt(sum(includedboth))
includedboth <- isincluded[2,]*isincluded[4,]
dichratios <- abs(EEGtimecourse7Hz[which(includedboth==1),2,])/abs(EEGtimecourse7Hz[which(includedboth==1),4,])
dichratios[which(dichratios==0)] <- 0.001
EEGmaskindex7[2,] <- colMeans(20*log10(dichratios),na.rm=TRUE)
EEGmaskindexSE7[2,] <- apply(20*log10(dichratios),2,sd,na.rm=TRUE)/sqrt(sum(includedboth))

@@

EEGmaskindex5[which(is.na(EEGmaskindex5))] <- 0
EEGmaskindex7[which(is.na(EEGmaskindex7))] <- 0
EEGmaskindexSE5[which(is.na(EEGmaskindexSE5))] <- 0
EEGmaskindexSE7[which(is.na(EEGmaskindexSE7))] <- 0

@@

if (plotsmoothed==0){
polygon(times[c(1:8000,8000:1)],pmin(c(EEGmeantime5Hz[1,]+EEGSEtime5Hz[1,],EEGmeantime5Hz[1,8000:1]-EEGSEtime5Hz[1,8000:1]),plotlims[4]),col=colpal[1],border=NA)
polygon(times[c(1:8000,8000:1)],c(EEGmeantime5Hz[3,]+EEGSEtime5Hz[3,],EEGmeantime5Hz[3,8000:1]-EEGSEtime5Hz[3,8000:1]),col=colpal[2],border=NA)
polygon(times[c(1:8000,8000:1)],c(EEGmeantime5Hz[4,]+EEGSEtime5Hz[4,],EEGmeantime5Hz[4,8000:1]-EEGSEtime5Hz[4,8000:1]),col=colpal[3],border=NA)

lines(times,pmin(EEGmeantime5Hz[1,],plotlims[4]),lwd=2,col=colpal[1])
# lines(times,EEGmeantime5Hz[2,],lwd=2,col='green')
lines(times,EEGmeantime5Hz[3,],lwd=2,col=colpal[2])
lines(times,EEGmeantime5Hz[4,],lwd=2,col=colpal[3])
}

if (plotsmoothed==1){
spl5aU <- smooth.spline(times,EEGmeantime5Hz[1,]+EEGSEtime5Hz[1,],df=20)
spl5aL <- smooth.spline(times,EEGmeantime5Hz[1,]-EEGSEtime5Hz[1,],df=20)
polygon(times[c(1:8000,8000:1)],pmin(c(spl5aU$y,spl5aL$y[8000:1]),plotlims[4]), col=colpal[1],border=NA)

@@

spl5a <- smooth.spline(times,pmin(EEGmeantime5Hz[1,],plotlims[4]),df=20)
lines(spl5a$x,spl5a$y,col=colpal[1],lwd=2)
# spl5b <- smooth.spline(times,EEGmeantime5Hz[2,],df=20)
# lines(spl5b$x,spl5b$y,col='green',lwd=2)
spl5c <- smooth.spline(times,EEGmeantime5Hz[3,],df=20)
lines(spl5c$x,spl5c$y,col=colpal[2],lwd=2)
spl5d <- smooth.spline(times,EEGmeantime5Hz[4,],df=20)
lines(spl5d$x,spl5d$y,col=colpal[3],lwd=2)

}
text(6,0.55,'5Hz',cex=2)
  legend(1.25,0.2,c('Target only','Monocular','Dichoptic'),lty=1,lwd=2,col=colpal[1:3],box.lwd=2)
dev.off()

@@

if (fixedphase==0){pdf('Figures/EEGdata.pdf', bg="transparent", height = 6, width = 15)}
if (fixedphase==1){pdf('Figures/EEGdataFP.pdf', bg="transparent", height = 6, width = 15)}
  par(mar=c(0.1,0.1,0.1,0.1))
  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,1), ylim=c(0,1))

@@

```{r EEGstats, include=FALSE, results='hide'}

@@

# calculate effect size

baseline <- rowMeans((20*log10(abs(EEGtimecourse5Hz[,1,1001:2000])/abs(EEGtimecourse5Hz[,3,1001:2000])) + 20*log10(abs(EEGtimecourse5Hz[,1,1001:2000])/abs(EEGtimecourse5Hz[,4,1001:2000])) + 20*log10(abs(EEGtimecourse7Hz[,2,1001:2000])/abs(EEGtimecourse7Hz[,3,1001:2000])) + 20*log10(abs(EEGtimecourse7Hz[,2,1001:2000])/abs(EEGtimecourse7Hz[,4,1001:2000])))/4)

laterepoch <- rowMeans((20*log10(abs(EEGtimecourse5Hz[,1,3001:4000])/abs(EEGtimecourse5Hz[,3,3001:4000])) + 20*log10(abs(EEGtimecourse5Hz[,1,3001:4000])/abs(EEGtimecourse5Hz[,4,3001:4000])) + 20*log10(abs(EEGtimecourse7Hz[,2,3001:4000])/abs(EEGtimecourse7Hz[,3,3001:4000])) + 20*log10(abs(EEGtimecourse7Hz[,2,3001:4000])/abs(EEGtimecourse7Hz[,4,3001:4000])))/4)

diff <- laterepoch-baseline

EEGeffectsize <- mean(diff)/sd(diff)

@@

```{r analyseMEGgroup, include=FALSE, results='hide'}

@@

if (fixedphase==0){            
        fspec1 <- NULL
        fspec2 <- NULL
        for (t in 1:(9009-binwidth)){
          spec <- fft(temp[t:(t+binwidth-1)])/binwidth
          fspec1[t] <- spec[tindex]
          fspec2[t] <- spec[mindex]
        }

        MEGtimecourse5Hz[s,cond,] <- fspec1
        MEGtimecourse7Hz[s,cond,] <- fspec2
}

if (fixedphase==1){  # used fixed phase dot product with sinewave
  # make a sine wave with the same phase as the EEG response
  spec <- fft(temp)/length(temp)
  synspec <- spec*0
  synspec[(targetF*8) + 1] <- spec[(targetF*8) + 1]
  sw1 <- Re(fft(synspec,inverse=TRUE))
  sw1 <- sw1/max(sw1)
  synspec <- spec*0
  synspec[(maskF*8) + 1] <- spec[(maskF*8) + 1]
  sw2 <- Re(fft(synspec,inverse=TRUE))
  sw2 <- sw2/max(sw2)

for (t in 1:(9000-binwidth)){
  MEGtimecourse5Hz[s,cond,t] <- mean(temp[t:(t+binwidth-1)] * sw1[t:(t+binwidth-1)])
  MEGtimecourse7Hz[s,cond,t] <- mean(temp[t:(t+binwidth-1)] * sw2[t:(t+binwidth-1)])
}}

@@

```{r plotMEGdata, include=FALSE, results='hide'}

@@

isincluded <- array(1, dim=c(4,dim(allspec)[1]))
if (fixedphase==0){
    for (cond in 1:4){
      # note that in allspec the DC component has been removed
      # so we don't need to add 1 to the index
      tempxy <- data.frame(Re(allspec[,cond,duration*targetF]),Im(allspec[,cond,duration*targetF]))
      D5 <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))
      tempxy <- data.frame(Re(allspec[,cond,duration*maskF]),Im(allspec[,cond,duration*maskF]))      
      D7 <- sqrt(mahalanobis(tempxy,colMeans(tempxy),cov(tempxy)))      
      isincluded[cond,which(D5>3)] <- 0
      isincluded[cond,which(D7>3)] <- 0
      for (s in 1:20){
        if (max(abs(MEGtimecourse5Hz[s,cond,]))>5){isincluded[cond,s] <- 0}
      }   
    }
  }

@@

    includedboth <- isincluded[1,]*isincluded[3,]
    monratios <- abs(MEGtimecourse5Hz[which(includedboth==1),1,])/abs(MEGtimecourse5Hz[which(includedboth==1),3,])
    MEGmaskindex5[1,] <- colMeans(20*log10(monratios),na.rm=TRUE)
    MEGmaskindexSE5[1,] <- apply(20*log10(monratios),2,sd,na.rm=TRUE)/sqrt(sum(includedboth))
    includedboth <- isincluded[1,]*isincluded[4,]
    dichratios <- abs(MEGtimecourse5Hz[which(includedboth==1),1,])/abs(MEGtimecourse5Hz[which(includedboth==1),4,])
    MEGmaskindex5[2,] <- colMeans(20*log10(dichratios),na.rm=TRUE)
    MEGmaskindexSE5[2,] <- apply(20*log10(dichratios),2,sd,na.rm=TRUE)/sqrt(sum(includedboth))

@@

MEGmaskindex5[which(is.na(MEGmaskindex5))] <- 0
MEGmaskindex7[which(is.na(MEGmaskindex7))] <- 0
MEGmaskindexSE5[which(is.na(MEGmaskindexSE5))] <- 0
MEGmaskindexSE7[which(is.na(MEGmaskindexSE7))] <- 0

@@

  }
  text(6,2.8,'5Hz',cex=2)
  legend(1.25,1,c('Target only', 'Monocular','Dichoptic'), lty=1,lwd=2,col=colpal[1:3],box.lwd=2)
  dev.off()

@@

if (fixedphase==0){pdf('Figures/MEGdata.pdf', bg="transparent", height = 6, width = 15)}
if (fixedphase==1){pdf('Figures/MEGdataFP.pdf', bg="transparent", height = 6, width = 15)}
  par(mar=c(0.1,0.1,0.1,0.1))
  plot(x=NULL,y=NULL,axes=FALSE, ann=FALSE, xlim=c(0,1), ylim=c(0,1))

@@

# calculate effect size

baseline <- rowMeans((20*log10(abs(MEGtimecourse5Hz[,1,1002:2003])/abs(MEGtimecourse5Hz[,3,1002:2003])) + 20*log10(abs(MEGtimecourse5Hz[,1,1002:2003])/abs(MEGtimecourse5Hz[,4,1002:2003])) + 20*log10(abs(MEGtimecourse7Hz[,2,1002:2003])/abs(MEGtimecourse7Hz[,3,1002:2003])) + 20*log10(abs(MEGtimecourse7Hz[,2,1002:2003])/abs(MEGtimecourse7Hz[,4,1002:2003])))/4)

laterepoch <- rowMeans((20*log10(abs(MEGtimecourse5Hz[,1,3004:4005])/abs(MEGtimecourse5Hz[,3,3004:4005])) + 20*log10(abs(MEGtimecourse5Hz[,1,3004:4005])/abs(MEGtimecourse5Hz[,4,3004:4005])) + 20*log10(abs(MEGtimecourse7Hz[,2,3004:4005])/abs(MEGtimecourse7Hz[,3,3004:4005])) + 20*log10(abs(MEGtimecourse7Hz[,2,3004:4005])/abs(MEGtimecourse7Hz[,4,3004:4005])))/4)

diff <- laterepoch-baseline

MEGeffectsize <- mean(diff)/sd(diff)

@@

```{r runIManalysis, include=FALSE, results='hide'}

samplerate <- 1000
targetF <- 7
maskF <- 5
tindex <- (10*(targetF+maskF)) + 1
duration <- 14
